Low Complexity Adaptive Beamformer\\ for Active Sonar Imaging

### 2015-10-14 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2015:10:14	23:05:15	(Wed14.Oct)

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This is because MVDR relies on estimating and inverting a covariance matrix. We have shown that in systems with less than $\approx$32 channels the complexity can be reduced and the implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014}. In larger systems the inversion step is the bottleneck and space reduction techniques more viable. In clinical ultrasound, where the transmit beams are focused and the spatial domain is sparse, we reduced  running our GPU implementation of MVDR in beamspace~\cite{Asen2013}. Two closely related techniques to recude is to form a simpler use principal component analysis (PCA)~\cite{Kim2014}, and to assume spatial stationarity to form more easily invertible Toepliz matrices~\cite{Asl2012}. At best these methods reduce the complexity to O($M^2$).

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in a system with focused transmit beams\todo{ha med at i US \o{}nsker man ulike vindus-typer p\aa{} forskjellig range?}. LCA applies a set of predefined windows, or tapers, and selects the one offering the best sidelobe suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a delay-and-sum (DAS) beamformer with an adaptive multi-taper extension. Such extensions up to 3 windows by e.g. Stankwitz in radar imaging~\cite{Stankwitz1995}, but unlike these LCA more windows, and add micro-steered variations of them.

>>>>

In many cases MVDR can improve the contrast and resolution of active sonar images compared to conventional static methods~\cite{Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}. However, while the computational complexity of conventional beamformers are linear with the number of channels, O($M$), MVDR is at O($M^3$). This is because MVDR relies on estimating and inverting a covariance matrix. We have shown that in systems with less than $\approx$32 channels the complexity can be reduced and the implementation accelerated significantly using graphics computing units (GPUs)~\cite{Buskenes2014}. For larger systems the inversion step is the bottleneck and space reduction techniques are desirable. In clinical ultrasound, where the transmit beams are focused, we processed our data in beamspace where the spatial sparsity allows much smaller matrices to be inverted~\cite{Asen2013}. Two closely related dimensionality reducing techniques used with MVDR includes using principal component analysis (PCA)~\cite{Kim2014}, and to assume spatial stationarity to form more easily invertible Toepliz matrices~\cite{Asl2012}. At best the mentioned methods are able to reduce the complexity to O($M^2$).

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams\todo{ha med at i US \o{}nsker man ulike vindus-typer p\aa{} forskjellig range?}. LCA applies a set of predefined windows, or tapers, and selects the one offering the best sidelobe suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a reduced window solution space. Alternatively, it may be viewed as a delay-and-sum (DAS) beamformer with an adaptive multi-taper extension. Multi-taper methods that select from up to 3 windows has been investigated by e.g. Stankwitz in radar imaging~\cite{Stankwitz1995}. LCA selects from even more windows, and also from micro-steered variations of them.
